### 2024-02-03 ms.qmd
Update ms.qmd
Updates

## Introduction
Behavioural flexibility, **the ability to adapt behaviour to changing environments** (**REFS**), is considered crucial for invasion success [@sol2002behavioural]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats (**REFS**) as behaving flexibly enhances their capacity to exploit new resources and face new challenges (**REFS**).

adaptive flexibility hypothesis says that flexibility may be relevant in early and late stages of invassion (Wright et al 2010). But at the same time behavioural flexibility  it is related to relative brain size (Sol et al 2002, 2008), that may be more fixed than we thought (controversy with Wright et al??). In contrast, other hyoptheses predict this is constant, or there is individual variation in beh flexibility (Wright et al 2010). Since some aspects of early environment can play a prominent role in shaping neural substrate, we hypothesize that beh flexibility may be affected by early environment. Also, since beh flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early developmental constraints. CORT can be essential in this process as it is associated to vertebrates response to environmental changes, and it is known to impact brain development (Zhu et al 2004). Also, temperature may be relevant, since it affects brain development in ectotherms, it can be one of the main environmental changes to face by animals, andvertebrates reponse to changes in temeprature is also associated with the production of CORT (Crino et al 2023).

Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context. To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe (**REFS**). Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_ (**REFS**), and are likely key for its invassive success (**REFS**). Our goal in this study was to investigate if behavioural flexibility is affected by the prenatal environment and the species specific response to it. Particularly, we were interested in the interactive effects of prenatal temperature and Corticosterone (CORT) - the main type of GC in reptiles [@Crino_2023] - because CORT is usually associated to vertebrates response to changes in temeprature [@Crino_2023], and both factors are known to impact brain development [@coomber_independent_1997; @zhu_prenatal_2004]. We hypothesized that, since behavioural flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early

@@

#### Experimental design
We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.  
To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A. domestica*) as a reward (see @fig-Methods). The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B). In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase. The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top. Lizards were tested once daily between 11 to 12 am, when lizards are active. The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days. We recorded and analysed manually all trials form the learning phases using a standard video player (IINA). PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses. We used as the main variable whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. Here, we analysed the data of the reversal as a measure of behavioural flexibility. As such, we only included in our analyses those individuals who passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.

@@

We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
**As such, we run a total of different 8 models employing Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). Each model consisted of four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations.'Choice', i.e. whether the individual chose correct (1) or not (0) was used as the response variable. The fixed effects of the model included a triple interaction between: trial ('Associative trial' or 'Reversal trial') as a numeric variable, and hormone treatmet ('CORT' versus 'Control') and the temperature at which eggs were incubated ('Cold' versus 'Hot') as factors. For the random effects, we employed lizard identity as a random intercept, and as a random slope we included the variable trial ('Associative' or 'Reversal') within each level of lizard identity. We used the resulting posterior of these models to evaluate learning differences between treatments within and between species and colour assigned. More specifically, we calculated learning slopes by using the estimates of the trial variable per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between 'Treatments' (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.**

@@

>>>>

## Introduction
Behavioural flexibility, the ability to adjust the behaviour in response to internal or external changes in an adaptive way by, e.g., problem-solving, innovation, or alterations in one behaviour in favour of another [@brown2010behavioral], is considered crucial for invasion success [@wright2010behavioral; @chapple2012can; @szabo2020can]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats, as behaving flexibly enhances their capacity to exploit new resources and face new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can]. Indeed, when closely related species of birds were compared, successful invaders showed a higher frequency of foraging innovations [@sol2000behavioural]. According to the adaptive flexibility hypothesis, behavioural flexibility can be more relevant during introduction, establishment, and spread of the population [@wright2010behavioral; @chapple2012can; @szabo2020can], stages that, nonetheless, can involve more than one generation [**REFS**]. Since behavioural flexibility can be affected by early environment [@jonson1976effect; @lui2017chronic], it is possible that the new prenatal environment faced by establishing populations can affect the behavioural flexibility of the new generations, with a consequent impact on invasion success. Therefore, we consider important to investigate the effect of developmental conditions on behavioural flexibility and whether their susceptibility to the early environment can predict the success of invasive species.  
Here, we are particularly interested in the interactive contribution of prenatal temperature and Glucocorticoids (GCs). There are several reasons why we believe these two aspects are salient for studying behavioural flexibility in the context of invasion ecology. First, GCs are the main group of hormones related to animals’ response towards alterations in homeostasis [@sapolsky_how_2000] that may become relevant when facing new environments. Furthermore, Gcs are one of the mechanisms mediating vertebrates’ response to changes in temperature that, at the same time, can have direct effects on animals' physiology and behaviour [@Crino_2023]. Secondly, there is evidence that both developmental GCs and thermal environment can impact cognition in different taxa [@jonson1976effect; @zhu_prenatal_2004; @amiel_hotter_2012; @lui2017chronic]. For instance, young Sprague-Dawley rats treated with Corticosterone (CORT) - the principal GC in rodents, birds, and reptiles [@Crino_2023] - performed worse than controls in a reversal learning task [@lui2017chronic], while exposure of pregnant guinea pigs to higher temperatures resulted in significant impariments with offspring behavoural flexibility [@jonson1976effect]. In addition, the cognitive effects of both factors are triggered by similar physiological mechanisms intertwined with brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004], an aspect that is considered essential in succesful invaders [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_smart_2011].

MODEL SPECIES AND POPULATIONS
To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe (**REFS**). Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_ (**REFS**), and are likely key for its invassive success (**REFS**).

HYPOTHESES
Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context.

@@

#### Experimental design
We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.  
To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A. domestica*) as a reward (see @fig-Methods). The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B). In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase. The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top. Lizards were tested once daily between 11 to 12 am, when lizards are active. The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days. We recorded and analysed manually all trials form the learning phases using a standard video player (IINA). PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses. We used as the main variable whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. Here, we analysed the data of the reversal as an index of behavioural flexibility [@brown2010behavioral].

@@

We performed the analyses for both species separately. As such, we fitted a total of two different Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). We had per each model four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations. We used 'Choice' - whether the individual chose correct (1) or not (0) - as the response variable, and trial, hormone ('CORT' versus 'Control'), incubation temperature ('Cold' versus 'Hot'), and their interaction as fixed factors. Lizard identity was employed as a random intercept, and as a random slope we included the trial within each level of lizard identity. We used the posterior distributions of these models to evaluate differences in learning rate between treatments and species. Learning slopes were obtained using the estimates of the variable trial per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between treatments (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) or species were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.

@@

#| label: tbl-data_2
